### 2019-11-25 TA_filter_despike.Rmd

---
title: "TA data filtration: despike()"
author: "samir"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  rmarkdown::html_document:
    theme: paper
    number_sections: false

---
<style type="text/css">

body{ /* Normal  */
      font-size: 18px;

}
</style>
```{r set-up, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
Sys.setlocale("LC_ALL", "en_US.UTF-8")
Sys.setenv(TZ='UTC') # on utilise UTC
rm(list = ls())
library(tidyverse)
library(robfilter)
library(seacarb)
library(gridExtra)
library(reshape2)
library(lubridate)
library(lmtest)
library(grid)
library(viridis)
library(dygraphs)
require("knitr")
library("lmodel2")
library(captioner)
library(xts)
library(seismicRoll)
library(scales)
library(knitr)

if (Sys.getenv("LOGNAME") == "gattuso") path = "../../pCloud\ Sync/Documents/experiments/exp168_awipev-CO2/"
if (Sys.getenv("LOGNAME") == "samir") path = "../../pCloud\ Sync/exp168_awipev-CO2/"
```

## Introduction

In this document, we try to deal with "cleaning TA dataset" because of high variability.
Variability of the dataset is partly due to the measurement of the seawater + acid during washing periode (noon and midnight).
We only deal with 2018-2019 TA data.

```{r prepare data, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
load(file = paste0(path,"fb_awipev-co2_server/ny-alesund/data/NRT_data/all_nydata_hour.Rdata"))
d_hour <- d_hour %>%
  dplyr::select(datetime, AT_filtered)

```

## Dataset

Here the TA dataset already cleaned by despike() at the end of the ny-alesund_cc.R script. If we zoom in November 2019 we can see the drop in data around midnight each day, due to acid washing cycle.

```{r plot data, echo=FALSE, warning=FALSE, message=FALSE}

# plot
at_fb_xts <- dplyr::select(d_hour, datetime,AT_filtered)
at_fb_xts <- as.xts(at_fb_xts, order.by = at_fb_xts$datetime)
dygraph(at_fb_xts, ylab="at") %>%
  dySeries("AT_filtered", color = "blue", strokeWidth = 0, label = "raw") %>%
  #dySeries("AT_filtered2", color = "red", strokeWidth = 0, label = "filt") %>%
  #dySeries("AT_filtered3", color = "green", strokeWidth = 0, label = "filt3") %>%
  dyHighlight(highlightCircleSize = 8,highlightSeriesBackgroundAlpha = 0.2,hideOnMouseOut = TRUE) %>%
  dyOptions(drawGrid = TRUE, drawPoints = TRUE, pointSize = 2,useDataTimezone = TRUE) %>%
  dyRangeSelector(height = 30)
```

## Despike() parameters settings
The idea was to find best parameters (n and k) in despike() in order to keep largest data (N = number of points) with a smallest standard deviation (V).

  - Create a dataframe with all posibilities for parameters n and k comprised between 0.1 - 1 for n and 25 - 221 for k.

  Final dataframe is created, compiling best n and K parameters: lowest standard deviation (V) and highest number of point (N).

Number of point > 80% and sd < 80 were choosen here.

## Despike() parameters results

```{r despike parameters, echo=FALSE, warning=FALSE, message=FALSE}

tmp <- NULL
#create a dataframe with all posibilities for parameters n and k comprised between 0.1 - 1 for n and 25 - 221 for k.
df <- expand.grid(n = seq(0.1, 1, length.out =0.1 ), k =as.integer(seq(25, 221, by = 4)))
df$N <- NULL
# create a tmp vector with different n, k parameters when i is different. i corresponds to the different lines of df and thus has different combinaisons of n,k.
for (i in 1:
     nrow(df)) {
  tmp <- despike(d_hour$AT_filtered, reference= "median", n=df$n[i], k=df$k[i], replace="NA")
  #create a new N column with the number of data. 
  df$N[i] <- length(tmp[!is.na(tmp)])
  #create a new V column with the standard deviation of each despike measurement.
  df$V[i] <- sd(tmp, na.rm = TRUE)
  }
summary(df)

#We want to keep data with highest N and smallest V
# We transform N in percentage for easy understanding.
df$percentN <- (df$N/max(df$N))*100

#final dataframe is created with bes despike() parameters for our arbitrary choice
acceptable <- df%>%
  mutate(N = ifelse(percentN > 70 & V <80, N, NA),
         k = ifelse(percentN > 70 & V <80, k, NA))%>%
  filter(!is.na(k))
kable(acceptable)

ggplot(data=df, aes(x = k, y = n)) +
  geom_point(aes(colour = N))

ggplot(data=df, aes(x = k, y = n)) +
  geom_point(aes(colour = V)) 
```

Best n and k were found for 0.2 > n < 0.3 and 109 >= K <= 221
```{r final plot, echo=FALSE, warning=FALSE, message=FALSE}
# final plot
at_fb_xts <- dplyr::select(d_hour, datetime,AT_filtered)
at_fb_xts <- as.xts(at_fb_xts, order.by = at_fb_xts$datetime)
dygraph(at_fb_xts, ylab="at") %>%
  dySeries("AT_filtered", color = "blue", strokeWidth = 0, label = "raw") %>%
  #dySeries("AT_filtered2", color = "red", strokeWidth = 0, label = "filt") %>%
  #dySeries("AT_filtered3", color = "green", strokeWidth = 0, label = "filt3") %>%
  dyHighlight(highlightCircleSize = 8,highlightSeriesBackgroundAlpha = 0.2,hideOnMouseOut = TRUE) %>%
  dyOptions(drawGrid = TRUE, drawPoints = TRUE, pointSize = 2,useDataTimezone = TRUE) %>%
  dyRangeSelector(height = 30)
```