### 2020-02-28 sal_temp.Rmd
New Rmd for salinity and temperature

---
title: "Salinity and temperature at Ny-Ålesund" 
author: "Jean-Pierre Gattuso and Samir Alliouane"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  rmarkdown::html_document:
    theme: paper
    number_sections: false

---
![ ](Images/AWIPEV-Logo.jpg){width=250px}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r Read and prepare ferrybox data, echo=FALSE, warning=FALSE, message=FALSE}

####define who is the user and define path
rm(list = ls())

if (Sys.getenv("LOGNAME") == "gattuso") path = "../../pCloud\ Sync/Documents/experiments/exp168_AWIPEV-CO2/fb_awipev-co2_server/ny-alesund/data/NRT_data/"

####Set environmental variables####
Sys.setenv(TZ="UTC")
if (!require("data.table")) install.packages("data.table")
##Provide in this section the parameters to be downloaded
#*********************************
#Use the following aggregation parameters
agg_time = "MINUTE" #SECOND, MINUTE, HOUR, DAY, MONTH, YEAR
#Calculate the following statistical values. 
agg_fun_1 = "MEAN" #MEDIAN
agg_fun_2 = "STDDEV"
agg_fun_3 = "N"
#*********************************

#### Download data ####
# Choose end date or take the date of today
# end_date <- ymd_hms("2017-12-31 23:59:59")
end_date <- ymd_hms(paste0(Sys.Date(), " 00:00:00 UTC"))
enddate <- format(end_date,"%Y-%m-%dT%H:%M:%S")

# Choose start date
start_date <- ymd_hms("2015-07-23 00:00:00")
startdate <- format(start_date, "%Y-%m-%dT%H:%M:%S")

# if enddate - startdate < 1 d one skips everything until the end of this script
if (end_date-start_date <= days(1)) {
  stop()
  }

if(agg_time=="MINUTE"){
  aggregate_string=paste0("&aggregate=",agg_time)
} else {
  aggregate_string=paste0("&aggregate=",agg_time,"&aggregateFunctions=",agg_fun_1,"&aggregateFunctions=",agg_fun_2,"&aggregateFunctions=",agg_fun_3)
}

code <- paste0("https://dashboard.awi.de/data-xxl/rest/data?beginDate=",startdate,"&endDate=",enddate,"&format=text/tab-separated-values",aggregate_string,
               "&sensors=station:svluwobs:fb_731101:sbe45_awi_0403:salinity",
               "&sensors=station:svluwobs:svluw2:sbe38_awi_0657:temperature")

data <- data.table::fread(code, encoding = "UTF-8", showProgress    = TRUE)
colnames(data) <- c("datetime", "Salinity", "Temperature" )
data$datetime <- ymd_hms(data$datetime)
```