### 2020-03-08 sal_temp.Rmd
sal_temp but not working yet

---
title: "Salinity and temperature at Ny-Ålesund" 
author: "Jean-Pierre Gattuso and Samir Alliouane"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  rmarkdown::html_document:
    theme: paper
    number_sections: false

---
![ ](Images/AWIPEV-Logo.jpg){width=250px}
```{r set-up, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
Sys.setlocale("LC_ALL", "en_US.UTF-8")
Sys.setenv(TZ='UTC') # on utilise UTC
rm(list = ls())
library(tidyverse)
library(robfilter)
library(seacarb)
library(gridExtra)
library(reshape2)
library(lubridate)
library(lmtest)
library(grid)
library(viridis)
library(dygraphs)
require("knitr")
library("lmodel2")
library(captioner)
library(xts)
library(seismicRoll)
library(scales)
if (!require("data.table")) install.packages("data.table")
knitr::opts_chunk$set(echo = TRUE)

fig_nums <- captioner()
table_nums <- captioner(prefix = "Table")

#define who is the user and define path
if (Sys.getenv("LOGNAME") == "gattuso") path = "../../pCloud\ Sync/Documents/experiments/exp168_awipev-CO2/"
if (Sys.getenv("LOGNAME") == "samir") path = "../../pCloud\ Sync/exp168_awipev-CO2/"

######## function to make regression plot with model I equation in title
ggreg <- function (fit, point_size=2) {
  ggplot(fit$model, aes_string(x = names(fit$model)[2],
                               y = names(fit$model)[1])) +
    geom_point(size = point_size, col = "blue") +
    stat_smooth(method = "lm", col = "black") +
    labs(title = paste(title, "\nAdj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                       "; Intercept =",signif(fit$coef[[1]],5 ),
                       "; Slope =",signif(fit$coef[[2]], 5),
                       "; P =",signif(summary(fit)$coef[2,4], 5))) +
    theme(plot.title = element_text(size=7))
}

#################### which.closest function
which.closest <- function(x, table, ...) {
  round(approx(x=table, y=1:length(table), xout=x, ...)$y)
}

#################### Regression function
# function regression plot with model II equation (MA) in title
## Dans labs ajout de la variable TITRE pour mettre titre avant chaque graphe
ggreg2 <- function (fit, xdata, ydata) { # x and y are the names of the variables
  fit_data <- data.frame(fit$x, fit$y)
  colnames(fit_data) = c(xdata, ydata)
reg <- fit$regression.results[2,] #one selects MA only
intercept <- reg$Intercept
slope <- reg$Slope
  ggplot(data = fit_data, aes_string(x = xdata, y = ydata)) + 
  geom_point(size = 2, col = "blue") +
  geom_abline(aes(intercept = fit$regression.results[2,2], slope = fit$regression.results[2,3]),
              colour = "blue")  + 
  labs(title = paste(titre,"\n Adj R2 = ", signif(fit$rsquare, 3),
                     "; Intercept =", signif(intercept, 3),
                     "; Slope =", signif(slope, 3),
                     "; P =", signif(fit$P.param, 3)))
}

mytheme <- theme_bw() +
  theme(axis.text.x=element_text(size=16, color="black"),
        axis.title.x=element_text(face="bold", size=16),
        axis.text.y=element_text(size=16, color="black"),
        axis.title.y=element_text(face="bold", size=16),
        plot.title = element_text(face="bold", size=14)
)

#################### Mytheme
Mytheme <- function(size_labs = 7, face_font="plain") {
  theme_bw() +
  theme(axis.text.x = element_text(face=face_font, size=size_labs, color="black"),
        axis.title.x = element_text(face=face_font, size=size_labs),
        axis.text.y = element_text(face=face_font, color="black", size=size_labs),
        axis.title.y = element_text(face=face_font, size=size_labs),
        axis.ticks.x = element_line(size=0.1),
        axis.ticks.y = element_line(size=0.1),
        axis.ticks.length = unit(1.1, "mm"),
        panel.grid.major = element_line(size = 0.25, color="black", linetype="dashed"),
        aspect.ratio = 1 / 2,
        plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "cm")
  )
}

ggplotRegression <- function(fit){
  require(ggplot2)
  ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red", se= FALSE) +
  labs(title = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     "Intercept =",signif(fit$coef[[1]],5 ),
                     " Slope =",signif(fit$coef[[2]], 5),
                     " P =",signif(summary(fit)$coef[2,4], 5)))
}
```

```{r Read and prepare ferrybox data, echo=FALSE, warning=FALSE, message=FALSE}

##Provide in this section the parameters to be downloaded
#*********************************
#Use the following aggregation parameters
agg_time = "HOUR" #SECOND, MINUTE, HOUR, DAY, MONTH, YEAR
#Calculate the following statistical values. 
agg_fun_1 = "MEAN" #MEDIAN
agg_fun_2 = "STDDEV"
agg_fun_3 = "N"
#*********************************

#### Download data ####
# Choose end date or take the date of today
 end_date <- ymd_hms("2019-12-31 23:59:59")
#end_date <- ymd_hms(paste0(Sys.Date(), " 00:00:00 UTC"))
enddate <- format(end_date,"%Y-%m-%dT%H:%M:%S")

# Choose start date
start_date <- ymd_hms("2012-01-01 00:00:00")
startdate <- format(start_date, "%Y-%m-%dT%H:%M:%S")

# if enddate - startdate < 1 d one skips everything until the end of this script
if (end_date-start_date <= days(1)) {
  stop()
  }

if(agg_time=="HOUR"){
  aggregate_string=paste0("&aggregate=",agg_time)
} else {
  aggregate_string=paste0("&aggregate=",agg_time,"&aggregateFunctions=",agg_fun_1,"&aggregateFunctions=",agg_fun_2,"&aggregateFunctions=",agg_fun_3)
}

code <- paste0("https://dashboard.awi.de/data-xxl/rest/data?beginDate=",startdate,"&endDate=",enddate,"&format=text/tab-separated-values",aggregate_string,
               "&sensors=station:svluwobs:fb_731101:sbe45_awi_0403:salinity",
               "&sensors=station:svluwobs:svluw2:sbe38_awi_0657:temperature")

data <- data.table::fread(code, encoding = "UTF-8", showProgress    = TRUE)
data <- tibble(data) %>%
  dplyr::rename(datetime = data$datetime,
                sal = $sal)
colnames(data) <- c("datetime", "sal", "temp" )
#data$datetime <- ymd_hms(data$datetime)
```

```{r average, fig.width=10, fig.height=4, echo=FALSE, message= FALSE}
data <- data %>%
  dplyr::mutate(hour = hour(data$datetime),
                date = 
                )

data_hour <- data %>%
  dplyr::group_by( date, hour) %>%
  dplyr::summarise(Salinity_filtered = mean(Salinity_filtered, na.rm = TRUE),
                   Temperature_filtered = mean(Temperature_filtered, na.rm = TRUE),
                   Temp_SBE38_filtered= mean(Temp_SBE38_filtered, na.rm = TRUE),
                   PCO2_Corr_filtered = mean(PCO2_Corr_filtered, na.rm = TRUE),
                   PeriodDeplpCO2 = mean(PeriodDeplpCO2, na.rm = TRUE),
                   PCO2_corr_contros_filtered= mean(PCO2_corr_contros_filtered, na.rm = TRUE),
                   AT_filtered= mean(AT_filtered, na.rm = TRUE),
                   HW_pH1_filtered= mean(HW_pH1_filtered, na.rm = TRUE),
                   HW_Temperature1_filtered= mean(HW_Temperature1_filtered, na.rm = TRUE),
                   T_seaF= mean(T_seaF, na.rm = TRUE),
                   phINT_filtered= mean(phINT_filtered, na.rm = TRUE),
                   phEXT_filtered= mean(phEXT_filtered, na.rm = TRUE),
                   voltINT= mean(voltINT, na.rm = TRUE),
                   voltEXT= mean(voltEXT, na.rm = TRUE),
                   pco2_inst= mean(pco2_inst, na.rm = TRUE),
                   ta_inst= mean(ta_inst, na.rm = TRUE),
                   seafet_inst= mean(seafet_inst, na.rm = TRUE) ) %>%
  dplyr::mutate(datetime = ymd_h(paste(date, hour, sep=" ", tz = "UTC"))) %>%
  dplyr::ungroup() %>% # this is to be able to perform the following changes
  dplyr::select(datetime, everything()) %>%
  dplyr::select(-c( hour)) %>% # not needed in the shiny display
  dplyr::arrange(desc(datetime))

```

```{r whole time series, fig.width=10, fig.height=4, echo=FALSE, message= FALSE}

#SAL
sal_xts <- dplyr::select(data, datetime, sal)
  #dplyr::filter(sal_fb >28)
sal_xts <- as.xts(sal_xts, order.by = sal_xts$datetime)
dygraph(sal_xts, group = "awipev", main="Salinity", ylab="Salinity") %>%
      dySeries("sal", color = "blue", strokeWidth = 0, label = "S") %>%
      dyHighlight(highlightCircleSize = 8,highlightSeriesBackgroundAlpha =0.2,hideOnMouseOut =TRUE) %>%
      dyOptions( drawGrid = TRUE, drawPoints = TRUE, pointSize = 2,useDataTimezone = TRUE) %>%
    dyLegend(show = "follow")   %>%
    dyAxis("y",valueRange = c(30, 37)) %>%
      dyRangeSelector(height = 30, dateWindow= NULL)

# temp
temp_xts <- dplyr::select(data, datetime, temp)
temp_xts <- as.xts(temp_xts, order.by = temp_xts$datetime)
dygraph(temp_xts, group = "awipev", main="Temperature", ylab="Temperature") %>%
      dySeries("temp_fb", color = "blue", strokeWidth = 0, label = "T") %>%
      dyAxis("y", valueRange = c(-2.5, 10)) %>%
      dyLegend(show = "follow")   %>%
      dyHighlight(highlightCircleSize = 8,highlightSeriesBackgroundAlpha = 0.5,hideOnMouseOut = TRUE) %>%
      dyOptions(drawGrid = TRUE, drawPoints = TRUE, pointSize = 2,useDataTimezone = TRUE) %>%
      dyRangeSelector(height = 30, dateWindow= NULL)
```