### 2020-03-27 paper.Rmd
revised tem_sal figs and regressions

---
title: "Paper on the carbonate chemistry at Ny-Ålesund" 
author: "Jean-Pierre Gattuso, Samir Alliouane and Philipp Fischer"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  rmarkdown::html_document:
    theme: paper
    number_sections: false
theme: cerulean
fig_width: 12 
fig_height: 8     
---

<style type="text/css">

body{ /* Normal  */
      font-size: 16px;
  }
td {  /* Table  */
  font-size: 14px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

```{r set-up, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
#R --quiet -e 'install.packages("plotly", type="binary", repos="http://cran.us.r-project.org")'
Sys.setlocale("LC_ALL", "en_US.UTF-8")
Sys.setenv(TZ='UTC') # on utilise UTC
rm(list = ls())
library(tidyverse)
library(seacarb)
library(gridExtra)
library(lubridate)
library(viridis)
require("knitr")
library(kableExtra)
library("lmodel2")
library(scales)
library(plotly)
library(htmlwidgets)
library(tsibble)
if (!require("directlabels")) install.packages("directlabels")
library("directlabels")
if (!require("pangaear")) install.packages("pangaea\r")
library("pangaear")
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = normalizePath("/Users/gattuso/not_synced_git/AWIPEV-CO2-git"))

#knitr::opts_chunk$set(fig.width=12, fig.height=8)

#fig_nums <- captioner()
#table_nums <- captioner(prefix = "Table")

#define who is the user and define paths
if (Sys.getenv("LOGNAME") == "gattuso") path_data = "../../pCloud\ Sync/Documents/experiments/exp168_awipev-CO2/"
if (Sys.getenv("LOGNAME") == "gattuso") path_fig = "/Users/gattuso/pCloud\ Sync/Documents/publications/publications_inpreparation/gattuso_etal_awipev-CO2/figures/"
if (Sys.getenv("LOGNAME") == "samir") path_data = "../../pCloud\ Sync/exp168_awipev-CO2/"

######## function to calculate 3 quantiles
qq <- function (data, digits) { 
  as.numeric(formatC(quantile(x = data, probs = c(0.025, 0.50, 0.975), na.rm = TRUE),
                     format = "f", digits = digits))
}

######## function to make regression plot with model I equation in title
ggreg <- function (fit, point_size=2) {
  ggplot(fit$model, aes_string(x = names(fit$model)[2],
                               y = names(fit$model)[1])) +
    geom_point(size = point_size, col = "blue") +
    stat_smooth(method = "lm", col = "black") +
    labs(title = paste(title, "\nAdj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                       "; Intercept =",signif(fit$coef[[1]],5 ),
                       "; Slope =",signif(fit$coef[[2]], 5),
                       "; P =",signif(summary(fit)$coef[2,4], 5))) +
    theme(plot.title = element_text(size=7))
}

#################### which.closest function
which.closest <- function(x, table, ...) {
  round(approx(x=table, y=1:length(table), xout=x, ...)$y)
}

#################### Regression function
# function regression plot with model II equation (MA) in title
## Dans labs ajout de la variable TITRE pour mettre titre avant chaque graphe
ggreg2 <- function (fit, xdata, ydata) { # x and y are the names of the variables
  fit_data <- data.frame(fit$x, fit$y)
  colnames(fit_data) = c(xdata, ydata)
reg <- fit$regression.results[2,] #one selects MA only
intercept <- reg$Intercept
slope <- reg$Slope
  ggplot(data = fit_data, aes_string(x = xdata, y = ydata)) + 
  geom_point(size = 2, col = "blue") +
  geom_abline(aes(intercept = fit$regression.results[2,2], slope = fit$regression.results[2,3]),
              colour = "blue")  + 
  labs(title = paste(titre,"\n Adj R2 = ", signif(fit$rsquare, 3),
                     "; Intercept =", signif(intercept, 3),
                     "; Slope =", signif(slope, 3),
                     "; P =", signif(fit$P.param, 3)))
}

mytheme <- theme_bw() +
  theme(axis.text.x=element_text(size=16, color="black"),
        axis.title.x=element_text(face="bold", size=16),
        axis.text.y=element_text(size=16, color="black"),
        axis.title.y=element_text(face="bold", size=16),
        plot.title = element_text(face="bold", size=14)
)

#################### Mytheme
Mytheme <- function(size_labs = 7, face_font="plain") {
  theme_bw() +
  theme(axis.text.x = element_text(face=face_font, size=size_labs, color="black"),
        axis.title.x = element_text(face=face_font, size=size_labs),
        axis.text.y = element_text(face=face_font, color="black", size=size_labs),
        axis.title.y = element_text(face=face_font, size=size_labs),
        axis.ticks.x = element_line(size=0.1),
        axis.ticks.y = element_line(size=0.1),
        axis.ticks.length = unit(1.1, "mm"),
        panel.grid.major = element_line(size = 0.25, color="black", linetype="dashed"),
        #aspect.ratio = 1 / 3,
        plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "cm")
  )
}

ggplotRegression <- function(fit){
ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red", se= FALSE) +
  labs(title = paste("Adj R2 = ", signif(summary(fit)$adj.r.squared, 5),
                     "Intercept =", signif(fit$coef[[1]],5 ),
                     " Slope =", signif(fit$coef[[2]], 5),
                     " P =", signif(summary(fit)$coef[2,4], 5)))
}
```

```{r read and clean data, echo=FALSE, warning=FALSE, message=FALSE}
d_all <- as_tibble(readRDS(file = paste0(path_data, "fb_data/d_all.rds"))) %>% 
  dplyr::filter(sal_fb > 28) %>% # remove salinity data lower than 28
  dplyr::mutate(year_month = format(datetime, '%Y-%m'))
```

## Whole time series

**To do**: Filter pH values at more than 8 m depth.

```{r whole time series, fig.width=10, fig.height=10, echo=FALSE, message= FALSE}
size_point <- 2 # size of data points
alpha <- 1 #transparency

#SAL SBE45
ts_sal <- d_all %>%
  ggplot(aes(x = datetime, y = sal_fb), na.rm=TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha=alpha, na.rm=TRUE, size=size_point) + 
  labs(title="Salinity FerryBox? SBE45",x="", y="Salinity") +
  Mytheme(size_labs = 8)
ts_sal_py <- ggplotly(ts_sal,dynamicTicks = TRUE)
filename.html <- paste0(path_fig, "ts_sal_py.html")
saveWidget(ts_sal_py, filename.html, selfcontained = T, libdir = "lib")

# TEMP ferrybox
ts_temp <- d_all %>%
  ggplot(aes(x = datetime, y = temp_fb), na.rm=TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha=alpha, na.rm=TRUE, size=size_point) + 
  labs(title="Temperature FerryBox",x="", y="In situ temperature (°C)") +
  Mytheme(size_labs = 8)

# PCO2 
ts_pCO2 <- d_all %>%
  ggplot() +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(aes(x = datetime, y = pco2_contros), 
             colour="blue", alpha=alpha, na.rm=TRUE, size=size_point) + 
  geom_point(aes(x = datetime, y = pco2calc), colour="red", alpha=alpha, na.rm=TRUE, size=size_point) + 
  labs(title="Mesured (blue) and calculated (CT+AT; red) pCO2",x="", y="pCO2 (uatm)") +
  Mytheme(size_labs = 8)

# SEAFET pH
ts_pH <- d_all %>%
  ggplot(aes(x = datetime, y = pHint_tot_sf), na.rm=TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha=alpha, na.rm=TRUE, size=size_point) + 
  labs(title="In situ pH",x="", y=expression(paste(pH[T]))) +
  Mytheme(size_labs = 8)

# TA
ts_at <- d_all %>%
  ggplot(aes(x = datetime, y = at_calc), na.rm=TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha=alpha, na.rm=TRUE, size=size_point) + 
  geom_point(aes(x = datetime, y = at), colour="red", alpha=alpha, na.rm=TRUE, size=size_point) + 
  labs(title="Total alkalinity calculated from salinity (blue) and discrete measurements (red)", x="", y="AT (umol/kg)") +
  Mytheme(size_labs = 8)

g <- cowplot::plot_grid(ts_sal, ts_temp, ts_pCO2, ts_pH, ts_at, align='v', ncol=1)
print(g)
```

Monthly distribution of seawater carbonate chemistry using a combination of a violin plot showing the relative frequency of the observations (shaded blue area) and a box plot showing the median, ﬁrst, and third quartiles, as well as outliers (blue).

```{r whole time series monthly box plots and stats, fig.width=10, fig.height=10, echo=FALSE, message= FALSE}
size_point <- 0.5 # size of data points
alpha <- 1 #transparency
ts_sal_bp <- d_all %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = sal_fb)) +
  geom_violin(fill='lightblue', alpha=0.5, size=0.2, trim = TRUE) +
  geom_boxplot(notch=TRUE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(y = "Salinity", x=NULL, title=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank())

ts_temp_bp <- d_all %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = temp_fb)) +
  geom_violin(fill='lightblue', alpha=0.5, size=0.2, trim = TRUE, na.rm=TRUE) +
  geom_boxplot(notch=TRUE, size=0.2, fill="grey50", outlier.color = "blue", 
               outlier.size = 0.3, na.rm=TRUE) +
  labs(y = "Temperature", x=NULL, title=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank())

ts_pCO2_bp <- d_all %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = pco2_contros)) +
  geom_violin(fill='lightblue', alpha=0.5, size=0.2, trim = TRUE, na.rm=TRUE) +
  geom_boxplot(notch=TRUE, size=0.2, fill="grey50", outlier.color = "blue", 
               outlier.size = 0.3, na.rm=TRUE) +
  labs(y = expression(paste(pCO[2])), x=NULL, title=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank())

ts_pH_bp <- d_all %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = pHint_tot_sf)) +
  geom_violin(fill='lightblue', alpha=0.5, size=0.2, trim = TRUE, na.rm=TRUE) +
  geom_boxplot(notch=TRUE, size=0.2, fill="grey50", outlier.color = "blue", 
               outlier.size = 0.3, na.rm=TRUE) +
  labs(y = expression(paste(pH[T])), x=NULL, title=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank())

ts_at_bp <- d_all %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = at_calc)) +
  geom_violin(fill='lightblue', alpha=0.5, size=0.2, trim = TRUE, na.rm=TRUE) +
  geom_boxplot(notch=TRUE, size=0.2, fill="grey50", outlier.color = "blue", 
               outlier.size = 0.3, na.rm=TRUE) +
  labs(y = "Total alkalinity", x=NULL, title=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank())

g <- cowplot::plot_grid(ts_sal_bp, ts_temp_bp, ts_pCO2_bp, ts_pH_bp, ts_at_bp, align='v', ncol=1)
print(g)

Salinity <- qq(d_all$sal_fb, digits=1)
Temperature <- qq(d_all$temp_fb, digits=1)
pCO2 <- qq(d_all$pco2_contros, digits=1)
pH <- qq(d_all$pHint_tot_sf, digits=3)
AT <- qq(d_all$at_calc, digits=1)
tab_quartiles <- as_tibble(rbind(Salinity, Temperature, pCO2, pH, AT)) %>%
  rename("2.5%"=V1, "50%"=V2, "97.5%"=V3) %>%
  mutate(Variable=c("Salinity", "Temperature", "pCO2", "pH", "AT")) %>%
  dplyr::select(Variable, everything())
kable(tab_quartiles, caption = "Key quantiles") %>%
  kable_styling(full_width=FALSE, bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

## Temperature vs salinity
We explore the use of a TS diagram to investigate changes in water sources. Monthly slopes of the T vs S plots are calculated. It seems that slopes are mostly not significant in winter and that they can be both positive and negative in summer.

**To do when the situation of the in situ CTDs will be clarified**: compare salinity values from the Ferrybox and in situ.

**To do**: add discrete pH values on plot.

```{r temperature vs salinity plots, echo=FALSE, message= FALSE, warning=FALSE, out.width="100%"}
# One tries to identify water masses coming from the Atlantic or freshwater
s <- c(seq(30.4, 35.3, length.out = 100)) # pCO2
t <- seq(-2, 8, length.out = 100) #temperature throughout the year
grid <- as_tibble(expand.grid(s=s, t=t))
grid$z <- rho(S = grid$s, T = grid$t, P = 1)

d <- d_all %>%
  dplyr::filter(year < 2020) %>%
  dplyr::mutate(month=month(datetime, label = TRUE, abbr = TRUE)) # abbreviated months rather than digital months
  #ggplot(aes(x = sal_fb, y = temp_insitu), na.rm=TRUE) +
temp_sal <-  ggplot() +
  geom_point(data = d, aes(x = sal_fb, y = temp_fb, fill=month), na.rm=TRUE,
             size=size_point, show.legend=FALSE) + 
  labs(title="Temperature vs salinity, both in FerryBox, blue lines show density", x="Salinity", y="Temp. (°C)") +
  facet_grid(year ~ month) +
  stat_smooth(data = d, aes(x = sal_fb, y = temp_fb, color=month),
              method=lm, na.rm=TRUE, se=FALSE, show.legend=FALSE) +
  #scale_color_viridis(name="Month", discrete=TRUE, option="plasma") +
  geom_contour(data = grid, aes(x = s, y = t, z = z), 
               stat = "contour", binwidth=1, show.legend=FALSE, na.rm=TRUE) +
  scale_x_continuous(limits = c(30.4, 35.3), expand = c(0,0) ) +
  scale_y_continuous(limits = c(-2,8.5), expand = c(0,0) ) +
  coord_fixed() +
  #annotate("text", label="Fresh and cold", x=31.5, y=-1.5) +
  #annotate("text", label="Salty and warm", x=34.5, y=8.1) +
  Mytheme(size_labs = 10)
print(temp_sal)
ggsave(temp_sal,filename = paste0(path_fig, "temp_sal.png"), width = 20, height = 20 ,units="cm")
temp_sal_py <- ggplotly(temp_sal,dynamicTicks = TRUE)
filename.html <- paste0(path_fig, "temp_sal_py.html")
saveWidget(temp_sal_py, filename.html, selfcontained = T, libdir = "lib")
```

```{r temperature vs salinity regressions, echo=FALSE, message= FALSE, warning=FALSE, out.width="50%"}
# AT as a function of salinity for every year
j <- 0
reg <- data.frame()
for (i in 1:length(unique(d$year_month))) { #one loops through all years
  j <- j + 1
  ym <- unique(d$year_month)[i]
  d_reg <- filter(d, year_month==ym)
  if (length(!is.na(d_reg$temp_fb)) > 2) { #there is no temp_insitu in 2016-08, hence skip. Not needed if one uses temp_fb
  lms <- summary(lm(data=d_reg, sal_fb ~ temp_fb))
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <- pf(lms$fstatistic[1],lms$fstatistic[2],lms$fstatistic[3],lower.tail=FALSE)
  reg <- rbind(reg, as.numeric(c(lms$coefficients[2,1], lms$coefficients[2,2], 
                                 lms$coefficients[2,4], lms$coefficients[1,1], lms$coefficients[1,2],
                                 lms$coefficients[1,4], lms$fstatistic[1], lms$fstatistic[3], lms$r.squared, prob)))
  }
}
names(reg) <- c("Slope", "SE_Slope", "P Slope","Intercept","SE int.","P int.", "F", "df",
                "R2","P_value")
row.names(reg) <- as.character(unique(d$year_month))
#transform in latex table and save
tab_reg_temp_sal <- reg
tmp_reg_temp_sal <- reg %>% # to enable plotting below
  dplyr::mutate(date = as.Date(paste0(row.names(reg), "-01")),
                month = month(date, label=TRUE, abbr = TRUE))
tab_reg_temp_sal <- tab_reg_temp_sal[sort(rownames(tab_reg_temp_sal)),] # order chronologically

#plot
fig_reg_temp_sal <- tmp_reg_temp_sal %>%
  dplyr::filter(P_value < .05) %>% #eliminate when correlation not significant
  ggplot() +
  labs(title = "Slope ± 2 x SE", subtitle = "3 months with correlation coefficient not significant were eliminated") +
  geom_point(aes(x=date, y=Slope, color=month, size=P_value)) +
  scale_color_viridis(discrete = TRUE) +
  #scale_fill_continuous(name="P value slope") +
  geom_hline(yintercept = 0) + 
  scale_size_continuous(range = c(3,7)) +
  geom_errorbar(aes(x=date, ymin=Slope-2*SE_Slope, ymax=Slope+2*SE_Slope), width=.2) +
  Mytheme(size_labs = 14) +
  guides(colour = guide_legend("Month", override.aes = list(size=4)), size = guide_legend("P value slope"))
kable(tab_reg_temp_sal, caption = "Regression lines", digits = c(3,3,5,1,3,3,1,0,2,5)) %>%
    kable_styling(full_width=FALSE, bootstrap_options = c("striped", "hover", "condensed", "responsive"))
print(fig_reg_temp_sal)
```

```{r density, fig.width=10, fig.height=10, echo=FALSE, message= FALSE}
# d_all <- d_all %>%
#   dplyr::mutate(den = den)
# ts_den <- d_all %>%
#   ggplot(aes(x = datetime, y = den), na.rm=TRUE) +
#   scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
#   geom_point(colour="black", na.rm=TRUE, size=size_point) + 
#   labs(title="Density",x="Time", y="Density") +
#   #coord_fixed() +
#   Mytheme(size_labs = 14)
# print(ts_den)
```

```{r pangaea data, fig.width=9, fig.height=6, echo=FALSE, message= FALSE}
# data_2015 <- pg_data(doi="10.1594/PANGAEA.896771", overwrite = TRUE)
# data_2016 <- pg_data(doi="10.1594/PANGAEA.896770", overwrite = TRUE)
# data_2017 <- pg_data(doi="10.1594/PANGAEA.896170", overwrite = TRUE)
# data_2018 <- pg_data(doi="10.1594/PANGAEA.897349", overwrite = TRUE)
# tmp <- rbind(data_2015[[1]]$data, data_2016[[1]]$data, data_2017[[1]]$data, data_2018[[1]]$data)
# tmp <- tmp[, c(1, 8, 11)] # select date time and salinity
# tmp$`Date/Time` <- as.POSIXct(strptime(tmp$`Date/Time`, "%Y-%m-%dT%H:%M", tz = "UTC"))
# 
# pg_ts <- tmp %>% #Pangaea time series
#   dplyr::filter(!are_duplicated(tmp)) %>%
#   dplyr::rename(datetime = `Date/Time`,
#                 sal = `Sal ([PSU])`,
#                 temp = `Temp [°C]`) %>%
#   as_tsibble(key = NULL, index = datetime) %>%
#   fill_gaps()
# 
# fig_sal_pangaea <- pg_ts %>%
#   ggplot() +
#   geom_point(aes(x=datetime, y=sal), na.rm = TRUE) +
#   labs(title="Salinity recovered from Pangaea") +
#   Mytheme(size_labs = 14)
# print(fig_sal_pangaea)
# 
# fig_temp_pangaea <- pg_ts %>%
#   ggplot() +
#   geom_point(aes(x=datetime, y=temp), na.rm = TRUE) +
#   labs(title="Temperature recovered from Pangaea") +
#   Mytheme(size_labs = 14)
# print(fig_temp_pangaea)
```

## Total alkalinity
It is proposed not to use the data from the Contros instrument. It is too unreliable and difficult to calibrate. We use instead the relationship between salinity and the discrete measurements of AT.

```{r Discrete salinity vs TA, echo=FALSE, warning=FALSE, message=FALSE, out.width="50%", fig.align="center"}
d_all_at_discrete <- d_all %>% #  To remove an outlier in red
  dplyr::select(datetime, month, closest_datetime, at, sal_fb)%>%
  dplyr::filter( !is.na(at) & !is.na(sal_fb) & sal_fb >30)

#fit <- lm(data = drop_na(d_all_at_discrete, at), at ~ sal)
fit <- lm(data = d_all_at_discrete, at ~ sal_fb)
slope_s_at <- fit$coefficients[[2]]
intercept_s_at <- fit$coefficients[[1]]
# at_s <- ggreg(fit)+ labs(x="Salinity", y="Discrete TA") + Mytheme()

at_s <- ggplot(d_all_at_discrete,aes(x=sal_fb, y= at, label = month))+
#  geom_text( aes(color=factor(month))) +
  geom_point(color = "blue") +
#  scale_color_discrete(guide="none")+
    labs(x="Salinity FerryBox", y="Discrete AT")  +
    stat_smooth(method = "lm", col = "black") +
    labs(title = paste("", "\nAdj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                       ";\nIntercept =",signif(fit$coef[[1]],5 ),
                       ";\nSlope =",signif(fit$coef[[2]], 5),
                       ";\nP =",signif(summary(fit)$coef[2,4], 5))) +
  theme_bw() +
  theme(aspect.ratio=1,
        axis.text=element_text(size=15),
        axis.title.y= element_text(size =15),
        axis.title.x= element_text(size =15),
        plot.margin = unit(c(0.5,1,1,1), "cm"),
        plot.title = element_text(size=15, hjust=0.05,margin = margin(l = 10, b = -70)))
print(at_s)

#ggsave(paste0("figures/Discrete_TA_salinity_AWIPEV.png"), at_s, width=15, height =11, units="cm")

# Salinity is a predictor of AT -> using the regression defined above
d_all$at_calc <- slope_s_at * d_all$sal_fb + intercept_s_at
```

The relationship between titrated total alkalinity and salinity is very good with an r2 of `r round(summary(fit)$r.squared, digits=2)`. Need to compare with other relationships from the literature.

## Consistency: pCO2

```{r pCO2, echo=FALSE, warning=FALSE, message=FALSE}
title <- "Contros vs calculated pCO2"
fit <- lmodel2(data = d_all,  pco2_contros ~ pco2calc , nperm = 99)

pco2calc_pCO2contros <-  ggplot(d_all, aes(x=pco2calc, y= pco2_contros))+
  geom_point(colour="blue", alpha = alpha)+
  geom_abline(aes(intercept = 0, slope = 1), linetype = "dashed") +
  geom_abline(aes(intercept = fit$regression.results[2,2], slope = fit$regression.results[2,3]),colour = "blue") +
  labs(title = paste(title,"\n Adj R2 = ", signif(fit$rsquare, 3), "\n",
                     "Intercept =", signif(fit$regression.results[2,2], 3),
                     "; Slope =", signif( fit$regression.results[2,3], 3),
                     "; P =", signif(fit$P.param, 3))) +
  theme(aspect.ratio=1, plot.title = element_text(size=7)) + 
  coord_fixed(ratio = 1, xlim=c(200,460) , ylim=c(200,460))+
  labs(x="Calculated pCO2", y="Measured pCO2") + theme_bw()  
print(pco2calc_pCO2contros)
ggsave(paste0("figures/pco2calc_pCO2contros.png"), pco2calc_pCO2contros, units="cm")

# average uncertainty calculated pCO2 assuming uncertainties of 4 umol/kg for AT and CT
e_pCO2 <- seacarb::errors(flag=15, 
                var1=1e-6*mean(d_all$at, na.rm=TRUE), var2=1e-6*mean(d_all$ct, na.rm=TRUE),
                S=mean(d_all$sal_fb, na.rm=TRUE), T=mean(d_all$temp_fb, na.rm=TRUE), P=0, Pt=0,
                Sit=0,k1k2="l", kf="dg", ks="d", pHscale="T", b="u74", 
                evar1=4e-6, evar2 = 4e-6)$pCO2
```

The relationship between the measured and calculated *p*CO~2~ is not very good. r2 (`r round(fit$rsquare, digits=2)`). Note that the accuracy of the pCO2 sensor is about 3 uatm (1% of reading) whereas it is `r round(e_pCO2, digits=1)` uatm. The later is estimated as described by Orr et al. (2019) assuming an uncertainty of 4 umol/kg for both AT and CT.