### 2020-11-13 awipev-CO2_PAR_sal.Rmd

---
title: "Carbonate chemistry at Ny-Ålesund: AWIPEV-CO~2~ Project" 
author: "Jean-Pierre Gattuso and Samir Alliouane"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  rmarkdown::html_document:
    theme: paper
    number_sections: false

---
![ ](Images/AWIPEV-Logo.jpg){width=250px}

```{r set-up, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
Sys.setlocale("LC_ALL", "en_US.UTF-8")

Sys.setenv(TZ='UTC') # on utilise UTC
rm(list = ls())
if (!require("tidyverse")) install.packages("tidyverse")
library(dplyr)
library(tidyverse)
if (!require("robfilter")) install.packages("robfilter")
if (!require("robfilter")) install.packages("robfilter")
library("robfilter")
if (!require("seacarb")) install.packages("seacarb")
library("seacarb")
if (!require("gridExtra")) install.packages("gridExtra")
library("gridExtra")
if (!require("reshape2")) install.packages("reshape2")
library("reshape2")
if (!require("lubridate")) install.packages("lubridate")
library("lubridate")
if (!require("lmtest")) install.packages("lmtest")
library("lmtest")
if (!require("grid")) install.packages("grid")
library(grid)
if (!require("viridis")) install.packages("viridis")
library("viridis")
if (!require("dygraphs")) install.packages("dygraphs")
library("dygraphs")
if (!require("knitr")) install.packages("knitr")
require("knitr")
if (!require("lmodel2")) install.packages("lmodel2")
library("lmodel2")
if (!require("captioner")) install.packages("captioner")
library("captioner")
if (!require("xts")) install.packages("xts")
library("xts")
if (!require("seismicRoll")) install.packages("seismicRoll")
library("seismicRoll")
if (!require("scales")) install.packages("scales")
library("scales")
if (!require("cowplot")) install.packages("cowplot")
library(cowplot)
if (!require("htmlwidgets")) install.packages("htmlwidgets")
library("htmlwidgets")
if (!require("hms")) install.packages("hms")
library("hms")
if (!require("ragg")) install.packages("hms")
library("ragg")
if (!require("skimr")) install.packages("skimr")
library("skimr")

knitr::opts_chunk$set(echo = TRUE)

fig_nums <- captioner()
table_nums <- captioner(prefix = "Table")

#define who is the user and define path
if (Sys.getenv("LOGNAME") == "gattuso") path = "../../pCloud\ Sync/Documents/experiments/exp168_awipev-CO2/"
if (Sys.getenv("LOGNAME") == "samir") path = "../../pCloud\ Sync/exp168_awipev-CO2/"

######## function to make regression plot with model I equation in title
ggreg <- function (fit, point_size=2) {
  ggplot(fit$model, aes_string(x = names(fit$model)[2],
                               y = names(fit$model)[1])) +
    geom_point(size = point_size, col = "blue") +
    stat_smooth(method = "lm", col = "black") +
    labs(title = paste(title, "\nAdj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                       "; Intercept =",signif(fit$coef[[1]],5 ),
                       "; Slope =",signif(fit$coef[[2]], 5),
                       "; P =",signif(summary(fit)$coef[2,4], 5))) +
    theme(plot.title = element_text(size=7))
}

#################### which.closest function
which.closest <- function(x, table, ...) {
  round(approx(x=table, y=1:length(table), xout=x, ...)$y)
}

#################### Regression function
# function regression plot with model II equation (MA) in title
## Dans labs ajout de la variable TITRE pour mettre titre avant chaque graphe
ggreg2 <- function (fit, xdata, ydata) { # x and y are the names of the variables
  fit_data <- data.frame(fit$x, fit$y)
  colnames(fit_data) = c(xdata, ydata)
reg <- fit$regression.results[2,] #one selects MA only
intercept <- reg$Intercept
slope <- reg$Slope
  ggplot(data = fit_data, aes_string(x = xdata, y = ydata)) + 
  geom_point(size = 2, col = "blue") +
  geom_abline(aes(intercept = fit$regression.results[2,2], slope = fit$regression.results[2,3]),
              colour = "blue")  + 
  labs(title = paste(titre,"\n Adj R2 = ", signif(fit$rsquare, 3),
                     "; Intercept =", signif(intercept, 3),
                     "; Slope =", signif(slope, 3),
                     "; P =", signif(fit$P.param, 3)))
}

#################### Function RMSE for model 2 regression root mean square error
lm2 <- function(x, y) { 
  fit <- lmodel2(y  ~ x , nperm = 99)
  intercept <- fit$regression.results[2,2]
  slope <- fit$regression.results[2,3]
  predicted <- intercept + x * slope
  error <- predicted - x
  my_list <- list(fit = fit, rmse = sqrt(mean(error^2, na.rm=TRUE)))
  return(my_list)
  }

#################### Mytheme
Mytheme <- function(size_labs = 7, face_font="plain") {
  theme_bw() +
  theme(axis.text.x = element_text(face=face_font, size=size_labs, color="black"),
        axis.title.x = element_text(face=face_font, size=size_labs),
        axis.text.y = element_text(face=face_font, color="black", size=size_labs),
        axis.title.y = element_text(face=face_font, size=size_labs),
        axis.ticks.x = element_line(size=0.1),
        axis.ticks.y = element_line(size=0.1),
        axis.ticks.length = unit(1.1, "mm"),
        panel.grid.major = element_line(size = 0.25, color="black", linetype="dashed"),
        aspect.ratio = 1 / 2,
        plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "cm")
  )
}

ggplotRegression <- function(fit){
ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red", se= FALSE) +
  labs(title = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     "Intercept =",signif(fit$coef[[1]],5 ),
                     " Slope =",signif(fit$coef[[2]], 5),
                     " P =",signif(summary(fit)$coef[2,4], 5)))
}

#################### MySkim
# myskim <- skim_with(numeric = sfl(min = ~min(., na.rm = TRUE),
#                                   max = ~max(., na.rm = TRUE)))
```

```{r Read, cleanup and prepare ferrybox data, echo=FALSE, warning=FALSE, message=FALSE}
d_hour <-
  readRDS(file = paste0(
    path,
    "fb_awipev-co2_server/ny-alesund/data/NRT_data/nydata_hour.rds"
  ))
d_hour <- d_hour %>%
  dplyr::rename(
    pco2_corr = pco2_corr_filtered,
    temp_dur = temp_dur_filtered,
    pco2_raw = pco2_raw_filtered,
    sal_fb = sal_fb_filtered,
    temp_insitu_11m = temp_insitu_11m_filtered,
    temp_fb = temp_fb_filtered,
    ph_dur = ph_dur_filtered
  ) %>% 
  arrange(datetime)
d_all <- as_tibble(d_hour) #d_all will be the merger
d_all <- d_all %>%
  dplyr::mutate_all(function(x) ifelse(is.nan(x), NA, x)) %>% # replace all NaN by NA
  dplyr::mutate(depth = ifelse( # create new variable depth
      pressure_insitu_ctd <= 2,
      1,
      ifelse(
        pressure_insitu_ctd > 2 & pressure_insitu_ctd <= 4,
        3,
        ifelse(
          pressure_insitu_ctd > 4 & pressure_insitu_ctd <= 6,
          5,
          ifelse(
            pressure_insitu_ctd > 6 & pressure_insitu_ctd <= 8,
            7,
            ifelse(pressure_insitu_ctd > 8, 9, NA)
          )
        )
      )
    )) %>%
  dplyr::rowwise() %>% 
  dplyr::mutate(n_depth_ctd = sum(is.na( # create new variable to identify the hours for which there are CTD data at more than 1 depth 
  c(
    sal_insitu_ctd_1m,
    sal_insitu_ctd_3m,
    sal_insitu_ctd_5m,
    sal_insitu_ctd_7m,
    sal_insitu_ctd_9m
  )
), na.rm = TRUE)) %>% 
  ungroup()
#restore datetime, date which are mangled by the mutates above, arrange ascending
d_all <- d_all %>% 
  dplyr::mutate(datetime = d_hour$datetime, 
                date = d_hour$date,
                monthd = lubridate::month(d_hour$datetime), # month as a number
                montht = lubridate::month(d_hour$datetime, label = TRUE, abbr = TRUE) # abbreviated months
  )
```

```{r Sal filter, fig.width=10, fig.height=4, echo=FALSE, message= FALSE}

########### Binding sal_fb + sal_insitu if we get gaps in sal_insitu = call this new column = sal_mix ########### 
d_all <- d_all %>%
  dplyr::mutate(sal_mix = ifelse(!is.na(sal_fb), sal_fb, sal_insitu_ctd_9m))
```

```{r data filter, echo=FALSE, fig.height=4, fig.width=10, message=FALSE, warning=FALSE}
#select mummer months
june2018 <- d_all %>%
  dplyr::select(datetime, par_insitu_profile_1m, par_insitu_profile_3m, par_insitu_profile_5m, par_insitu_profile_7m, par_insitu_profile_9m, par_air, turb_fb, sal_fb, sal_insitu_ctd_1m, sal_insitu_ctd_3m, sal_insitu_ctd_5m, sal_insitu_ctd_7m, sal_insitu_ctd_9m) %>% 
  dplyr::filter(datetime >= "2018-06-01 00:00:00" & datetime <= "2018-06-30 23:59:59")%>% 
  dplyr::mutate(day = day(datetime))

july2018 <- d_all %>%
  dplyr::select(datetime, par_insitu_profile_1m, par_insitu_profile_3m, par_insitu_profile_5m, par_insitu_profile_7m, par_insitu_profile_9m, par_air, turb_fb, sal_fb, sal_insitu_ctd_1m, sal_insitu_ctd_3m, sal_insitu_ctd_5m, sal_insitu_ctd_7m, sal_insitu_ctd_9m) %>% 
  dplyr::filter(datetime >= "2018-07-01 00:00:00" & datetime <= "2018-07-31 23:59:59")%>% 
  dplyr::mutate(day = day(datetime))

august2018 <- d_all %>%
  dplyr::select(datetime, par_insitu_profile_1m, par_insitu_profile_3m, par_insitu_profile_5m, par_insitu_profile_7m, par_insitu_profile_9m, par_air, turb_fb, sal_fb, sal_insitu_ctd_1m, sal_insitu_ctd_3m, sal_insitu_ctd_5m, sal_insitu_ctd_7m, sal_insitu_ctd_9m) %>% 
  dplyr::filter(datetime >= "2018-08-01 00:00:00" & datetime <= "2018-08-31 23:59:59") %>% 
  dplyr::mutate(day = day(datetime))

# summarise the max PAR/sal by day
sumjune2018 <- june2018 %>% 
  dplyr::group_by(day) %>% 
  dplyr:: summarise(maxPAR1m = max(par_insitu_profile_1m, na.rm=TRUE),
                  maxPAR3m = max(par_insitu_profile_3m, na.rm=TRUE),
                  maxPAR5m = max(par_insitu_profile_5m, na.rm=TRUE),
                  maxPAR7m = max(par_insitu_profile_7m, na.rm=TRUE),
                  maxPAR9m = max(par_insitu_profile_9m, na.rm=TRUE),
                  maxSAL1m = max(sal_insitu_ctd_1m, na.rm=TRUE),
                  maxSAL3m = max(sal_insitu_ctd_3m, na.rm=TRUE),
                  maxSAL5m = max(sal_insitu_ctd_5m, na.rm=TRUE),
                  maxSAL7m = max(sal_insitu_ctd_7m, na.rm=TRUE),
                  maxSAL9m = max(sal_insitu_ctd_9m, na.rm=TRUE),
                  maxPARair = max(par_air, na.rm=TRUE),
                  nb =n())

sumjuly2018 <- july2018 %>% 
  dplyr::group_by(day) %>% 
  dplyr:: summarise(maxPAR1m = max(par_insitu_profile_1m, na.rm=TRUE),
                  maxPAR3m = max(par_insitu_profile_3m, na.rm=TRUE),
                  maxPAR5m = max(par_insitu_profile_5m, na.rm=TRUE),
                  maxPAR7m = max(par_insitu_profile_7m, na.rm=TRUE),
                  maxPAR9m = max(par_insitu_profile_9m, na.rm=TRUE),
                  maxSAL1m = max(sal_insitu_ctd_1m, na.rm=TRUE),
                  maxSAL3m = max(sal_insitu_ctd_3m, na.rm=TRUE),
                  maxSAL5m = max(sal_insitu_ctd_5m, na.rm=TRUE),
                  maxSAL7m = max(sal_insitu_ctd_7m, na.rm=TRUE),
                  maxSAL9m = max(sal_insitu_ctd_9m, na.rm=TRUE),
                  maxPARair = max(par_air, na.rm=TRUE),
                  nb =n())

sumaugust2018 <- august2018 %>% 
  dplyr::group_by(day) %>% 
  dplyr:: summarise(maxPAR1m = max(par_insitu_profile_1m, na.rm=TRUE),
                  maxPAR3m = max(par_insitu_profile_3m, na.rm=TRUE),
                  maxPAR5m = max(par_insitu_profile_5m, na.rm=TRUE),
                  maxPAR7m = max(par_insitu_profile_7m, na.rm=TRUE),
                  maxPAR9m = max(par_insitu_profile_9m, na.rm=TRUE),
                  maxSAL1m = max(sal_insitu_ctd_1m, na.rm=TRUE),
                  maxSAL3m = max(sal_insitu_ctd_3m, na.rm=TRUE),
                  maxSAL5m = max(sal_insitu_ctd_5m, na.rm=TRUE),
                  maxSAL7m = max(sal_insitu_ctd_7m, na.rm=TRUE),
                  maxSAL9m = max(sal_insitu_ctd_9m, na.rm=TRUE),
                  maxPARair = max(par_air, na.rm=TRUE),
                  nb =n())
# tables
kable(sumjune2018, caption ="June 2018")
kable(sumjuly2018, caption ="July 2018")
kable(sumaugust2018, caption ="August 2018")
```
### Plot

```{r PAR / SAL Dygraph, fig.width=10, fig.height=4, echo=FALSE, message= FALSE}
#xts + dygraph
sal_xts <- dplyr::select(d_all, datetime,sal_insitu_ctd_1m, sal_insitu_ctd_3m, sal_insitu_ctd_5m, sal_insitu_ctd_7m, sal_insitu_ctd_9m, par_insitu_profile_1m, par_insitu_profile_3m, par_insitu_profile_5m, par_insitu_profile_7m, par_insitu_profile_9m, par_air)%>%
   dplyr::filter(!is.na(datetime))

sal_xts <- as.xts(sal_xts, order.by = sal_xts$datetime)
dygraph(sal_xts, group = "awipev", main="Salinity and PAR (in field)", ylab="") %>%
   dySeries("sal_insitu_ctd_1m", color = "blue", strokeWidth = 0, label = "SAL 1 m") %>%
   dySeries("sal_insitu_ctd_3m", color = "blue", strokeWidth = 0, label = "SAL 3 m") %>%
   dySeries("sal_insitu_ctd_5m", color = "blue", strokeWidth = 0, label = "SAL 5 m") %>%
   dySeries("sal_insitu_ctd_7m", color = "blue", strokeWidth = 0, label = "SAL 7 m") %>%
   dySeries("sal_insitu_ctd_9m", color = "blue", strokeWidth = 0, label = "SAL 9 m") %>%
   dySeries("par_insitu_profile_1m", color = "red", strokeWidth = 0, label = "PAR 1 m",axis = "y2") %>%
   dySeries("par_insitu_profile_3m", color = "red", strokeWidth = 0, label = "PAR 3 m",axis = "y2") %>%
   dySeries("par_insitu_profile_5m", color = "red", strokeWidth = 0, label = "PAR 5 m",axis = "y2") %>%
   dySeries("par_insitu_profile_7m", color = "red", strokeWidth = 0, label = "PAR 7 m",axis = "y2") %>%
   dySeries("par_insitu_profile_9m", color = "red", strokeWidth = 0, label = "PAR 9 m",axis = "y2") %>%
     dySeries("par_air", color = "orange", strokeWidth = 0, label = "PAR AIR",axis = "y2") %>%
dyHighlight(highlightCircleSize = 8,highlightSeriesBackgroundAlpha =0.2,hideOnMouseOut =TRUE) %>%
      dyOptions( drawGrid = TRUE, drawPoints = TRUE, pointSize = 2,useDataTimezone = TRUE) %>%
    dyLegend(show = "follow")   %>%
   dyAxis("y2",valueRange = c(0, 1350), label = "PAR") %>%
      dyRangeSelector(height = 30, dateWindow= NULL)

#plot 2
#Delete the NA line with datetime
fb_xts <- dplyr::select(d_all, datetime,sal_fb, turb_fb)
fb_xts <- fb_xts[-which(is.na(fb_xts$datetime)),]
fb_xts <- as.xts(fb_xts, order.by = fb_xts$datetime)
dygraph(fb_xts, group = "awipev", main="Salinity and Turbidity (in FerryBox)", ylab="") %>%
   dySeries("sal_fb", color = "blue", strokeWidth = 0, label = "SAL FB") %>%
   dySeries("turb_fb", color = "brown", strokeWidth = 0, pointSize = 4,label = "Turb FB",axis = "y2") %>%
dyHighlight(highlightCircleSize = 8,highlightSeriesBackgroundAlpha =0.2,hideOnMouseOut =TRUE) %>%
      dyOptions( drawGrid = TRUE, drawPoints = TRUE, pointSize = 2,useDataTimezone = TRUE) %>%
    dyLegend(show = "follow")   %>%
  dyAxis("y2",valueRange = c(0, 1000), label = "PAR") %>%
      dyRangeSelector(height = 30, dateWindow= NULL)
```

## Relationships by depths (first test)

```{r PAR / SAL relationship, echo=FALSE, fig.asp=0.5, out.width = "100%", message=FALSE, warning=FALSE}
#1m
fit <- lm(data = d_all, par_insitu_profile_1m ~ sal_insitu_ctd_1m)
p1 <- ggplot(d_all,aes(x=sal_insitu_ctd_1m, y= par_insitu_profile_1m, label = monthd))+
  geom_text( aes(color=factor(monthd))) +
  scale_color_discrete(guide="none")+
    labs(x="Salinity CTD 1 m", y="PAR Profile 1 m", color = "Months")  +
    stat_smooth(method = "lm", col = "black") +
    labs(title = paste("", "\nAdj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                       ";\nIntercept =",signif(fit$coef[[1]],5 ),
                       ";\nSlope =",signif(fit$coef[[2]], 5),
                       ";\nP =", signif(summary(fit)$coef[2,4], 5),
                       ";\nRMSE =", signif(sqrt(mean(fit$residuals^2)), 5))) +
  theme_bw() + theme(aspect.ratio=1,
        axis.text=element_text(size=7),
        axis.title.y= element_text(size =7),
        axis.title.x= element_text(size =7),
        plot.margin = unit(c(0.5,1,1,1), "cm"),
        plot.title = element_text(size=7, hjust=0.05,margin = margin(l = 10, b = -70)))

#3m
fit <- lm(data = d_all, par_insitu_profile_3m ~ sal_insitu_ctd_3m)
p3 <- ggplot(d_all,aes(x=sal_insitu_ctd_3m, y= par_insitu_profile_3m, label = monthd))+
  geom_text( aes(color=factor(monthd))) +
  scale_color_discrete(guide="none")+
    labs(x="Salinity CTD 3 m", y="PAR Profile 3 m", color = "Months")  +
    stat_smooth(method = "lm", col = "black") +
    labs(title = paste("", "\nAdj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                       ";\nIntercept =",signif(fit$coef[[1]],5 ),
                       ";\nSlope =",signif(fit$coef[[2]], 5),
                       ";\nP =", signif(summary(fit)$coef[2,4], 5),
                       ";\nRMSE =", signif(sqrt(mean(fit$residuals^2)), 5))) +
  theme_bw() + theme(aspect.ratio=1,
        axis.text=element_text(size=7),
        axis.title.y= element_text(size =7),
        axis.title.x= element_text(size =7),
        plot.margin = unit(c(0.5,1,1,1), "cm"),
        plot.title = element_text(size=7, hjust=0.05,margin = margin(l = 10, b = -70)))
#5m
fit <- lm(data = d_all, par_insitu_profile_5m ~ sal_insitu_ctd_5m)
p5 <- ggplot(d_all,aes(x=sal_insitu_ctd_5m, y= par_insitu_profile_5m, label = monthd))+
  geom_text( aes(color=factor(monthd))) +
  scale_color_discrete(guide="none")+
    labs(x="Salinity CTD 5 m", y="PAR Profile 5 m", color = "Months")  +
    stat_smooth(method = "lm", col = "black") +
    labs(title = paste("", "\nAdj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                       ";\nIntercept =",signif(fit$coef[[1]],5 ),
                       ";\nSlope =",signif(fit$coef[[2]], 5),
                       ";\nP =", signif(summary(fit)$coef[2,4], 5),
                       ";\nRMSE =", signif(sqrt(mean(fit$residuals^2)), 5))) +
  theme_bw() + theme(aspect.ratio=1,
        axis.text=element_text(size=7),
        axis.title.y= element_text(size =7),
        axis.title.x= element_text(size =7),
        plot.margin = unit(c(0.5,1,1,1), "cm"),
        plot.title = element_text(size=7, hjust=0.05,margin = margin(l = 10, b = -70)))
#7m
fit <- lm(data = d_all, par_insitu_profile_7m ~ sal_insitu_ctd_7m)
p7 <- ggplot(d_all,aes(x=sal_insitu_ctd_7m, y= par_insitu_profile_7m, label = monthd))+
  geom_text( aes(color=factor(monthd))) +
  scale_color_discrete(guide="none")+
    labs(x="Salinity CTD 7 m", y="PAR Profile 7 m", color = "Months")  +
    stat_smooth(method = "lm", col = "black") +
    labs(title = paste("", "\nAdj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                       ";\nIntercept =",signif(fit$coef[[1]],5 ),
                       ";\nSlope =",signif(fit$coef[[2]], 5),
                       ";\nP =", signif(summary(fit)$coef[2,4], 5),
                       ";\nRMSE =", signif(sqrt(mean(fit$residuals^2)), 5))) +
  theme_bw() + theme(aspect.ratio=1,
        axis.text=element_text(size=7),
        axis.title.y= element_text(size =7),
        axis.title.x= element_text(size =7),
        plot.margin = unit(c(0.5,1,1,1), "cm"),
        plot.title = element_text(size=7, hjust=0.05,margin = margin(l = 10, b = -70)))
#9m
fit <- lm(data = d_all, par_insitu_profile_9m ~ sal_insitu_ctd_9m)
p9 <- ggplot(d_all,aes(x=sal_insitu_ctd_9m, y= par_insitu_profile_9m, label = monthd))+
  geom_text( aes(color=factor(monthd))) +
  scale_color_discrete(guide="none")+
    labs(x="Salinity CTD 9 m", y="PAR Profile 9 m", color = "Months")  +
    stat_smooth(method = "lm", col = "black") +
    labs(title = paste("", "\nAdj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                       ";\nIntercept =",signif(fit$coef[[1]],5 ),
                       ";\nSlope =",signif(fit$coef[[2]], 5),
                       ";\nP =", signif(summary(fit)$coef[2,4], 5),
                       ";\nRMSE =", signif(sqrt(mean(fit$residuals^2)), 5))) +
  theme_bw() + theme(aspect.ratio=1,
        axis.text=element_text(size=7),
        axis.title.y= element_text(size =7),
        axis.title.x= element_text(size =7),
        plot.margin = unit(c(0.5,1,1,1), "cm"),
        plot.title = element_text(size=7, hjust=0.05,margin = margin(l = 10, b = -70)))

print(p1)
print(p3)
print(p5)
print(p7)
print(p9) 
```

## Relationships by summer months (second test)

```{r PAR / SAL relationship summer months, echo=FALSE, fig.asp=0.5, out.width = "100%", message=FALSE, warning=FALSE}
#june
fit <- lm(data = sumjune2018, maxPAR7m ~ maxSAL7m)
p_june <- ggplot(sumjune2018,aes(x=maxSAL7m, y= maxPAR7m, label = day))+
  geom_text( aes(color=factor(day))) +
  scale_color_discrete(guide="none")+
    labs(x="Salinity June 7 m", y="PAR June 7 m", color = "Days")  +
    stat_smooth(method = "lm", col = "black") +
    labs(title = paste("", "\nAdj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                       ";\nIntercept =",signif(fit$coef[[1]],5 ),
                       ";\nSlope =",signif(fit$coef[[2]], 5),
                       ";\nP =", signif(summary(fit)$coef[2,4], 5),
                       ";\nRMSE =", signif(sqrt(mean(fit$residuals^2)), 5))) +
  ylim(0,1500) +
  theme_bw() + theme(aspect.ratio=1,
        axis.text=element_text(size=7),
        axis.title.y= element_text(size =7),
        axis.title.x= element_text(size =7),
        plot.margin = unit(c(0.5,1,1,1), "cm"),
        plot.title = element_text(size=7, hjust=0.05,margin = margin(l = 10, b = -70)))
#july
fit <- lm(data = sumjuly2018, maxPAR7m ~ maxSAL7m)
p_july <- ggplot(sumjuly2018,aes(x=maxSAL7m, y= maxPAR7m, label = day))+
  geom_text( aes(color=factor(day))) +
  scale_color_discrete(guide="none")+
    labs(x="Salinity July 7 m", y="PAR July 7 m", color = "Days")  +
    stat_smooth(method = "lm", col = "black") +
    labs(title = paste("", "\nAdj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                       ";\nIntercept =",signif(fit$coef[[1]],5 ),
                       ";\nSlope =",signif(fit$coef[[2]], 5),
                       ";\nP =", signif(summary(fit)$coef[2,4], 5),
                       ";\nRMSE =", signif(sqrt(mean(fit$residuals^2)), 5))) +
  ylim(0,1500) +
  theme_bw() + theme(aspect.ratio=1,
        axis.text=element_text(size=7),
        axis.title.y= element_text(size =7),
        axis.title.x= element_text(size =7),
        plot.margin = unit(c(0.5,1,1,1), "cm"),
        plot.title = element_text(size=7, hjust=0.05,margin = margin(l = 10, b = -70)))

#august
fit <- lm(data = sumaugust2018, maxPAR7m ~ maxSAL7m)
p_august <- ggplot(sumaugust2018,aes(x=maxSAL7m, y= maxPAR7m, label = day))+
  geom_text( aes(color=factor(day))) +
  scale_color_discrete(guide="none")+
    labs(x="Salinity August 7 m", y="PAR August 7 m", color = "Days")  +
    stat_smooth(method = "lm", col = "black") +
    labs(main ="August 2018", title = paste("", "\nAdj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                       ";\nIntercept =",signif(fit$coef[[1]],5 ),
                       ";\nSlope =",signif(fit$coef[[2]], 5),
                       ";\nP =", signif(summary(fit)$coef[2,4], 5),
                       ";\nRMSE =", signif(sqrt(mean(fit$residuals^2)), 5))) +
  ylim(0,1500) +
  theme_bw() + theme(aspect.ratio=1,
        axis.text=element_text(size=7),
        axis.title.y= element_text(size =7),
        axis.title.x= element_text(size =7),
        plot.margin = unit(c(0.5,1,1,1), "cm"),
        plot.title = element_text(size=7, hjust=0.05,margin = margin(l = 10, b = -70)))

print(p_june)
print(p_july)
print(p_august)
```