### 2022-03-30 awipev-CO2_presentation.Rmd
Two columns

---
title: "**First high-frequency time series of carbonate chemistry in the Arctic**" 
author: "**Jean-Pierre Gattuso**^1,2^ and Samir Alliouane^1^ and Philipp Fisher^3^"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  html_document:
    # code_folding: show
#    css: acid.css
    fig_caption: yes
    fig_width: 2
    fig_height: 2
    highlight: default    
    number_sections: no
    theme: yeti
    toc: no
    toc_float: no
fontsize: 32pt
language: Australian
#css: acid.css
## mainfont: PT Serif  
## monofont: PT Mono
#csl: frontiers.csl
## sansfont: PT Sans
#bibliography: ArcticKelp.bib
## papersize: A4
---
<!-- <style type="text/css"> -->
<!-- .main-container { -->
<!--   max-width: 100% !important; -->
<!--   margin: auto; -->
<!-- } -->
</style>
<style type="text/css">
.main-container {
  max-width: 100% !important;
  margin: auto;
}
body{
  font-size: 24pt;
}
h1,h2,h3,h4,h5,h6{
  font-size: 32pt;
}
</style>

^1^Laboratoire d'Océanographie de Villefranche, CNRS-Sorbonne Université\
^2^Iddri, Sciences Po\
^3^Alfred Wegener Institute for Polar and Marine Research\

```{r set-up, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
  #if (Sys.getenv("LOGNAME") == "gattuso") path = "."
```

![](images/AWIPEV-Logo.jpg){width="250px"}

***
# **Introduction**

- CO<sub>2</sub> is an acid gas (it produces a weak acid when combined with water)
- Each of us adds 4 kg CO<sub>2</sub> per day to the ocean (increasing acidity, reducing pH)
- Arctic surface water will become corrosive for aragonite: 50% in 2050, 100% in 2100
- Critical to document the changes but no high frequency time-series

***

Fransner et al. (2022)
![ ](images/map_fransner.png){width="50%"}
![ ](images/pH_fransner.png)

***

# **Instruments installed**

- Water pumped from 12 m water depth at Ny-Ålesund
- **July 2015**: Two pCO<sub>2</sub> sensors rotating every year (in FerryBox, 1 min) 
- **February 2016**: Total alkalinity analyser (in FerryBox, 90 min) but lots of technical issues
- For calibration and quality insurance: weekly discrete samples for high-precision measurements of dissolved inorganic carbon and total alkalinity
- **August 2017**: pH measurements
  - Continuous measurements in fjord water (seaFET)
  - DURAFET pH sensor in the FerryBox
  - Discrete pH samples once a month for calibration
- Regular maintenance twice a year (except during the Covid-19 pandemic)
- Came to an end in December 2021

## **Instruments in the FerryBox**
![ ](images/container.png)

![ ](images/intake.png)

![ ](images/ferrybox.jpg)

## seaFET pH on the REMOS
![ ](images/seafet.png)

***

## No longer in real-time: https://awipev-co2.obs-vlfr.fr
![ ](images/web_page.png)

***

## **Availability of data**
:::: {style="display: flex;"}
::: {class='col-md-4'}
Many challenges:

- extreme environmental conditions
- frozen pipes
- damages resulting from icebergs pounding on the field instruments
- weeks to months waiting for warmer temperatures to deice or to fly technical staff, including divers
- No access for extended periods of time during the Covid-19 pandemic

Nevertheless:

- data usable 50 to 76% of the time during the period of measurement
- pCO<sub>2</sub> and pH data are available throughout the year and well distributed across months, including winter months
:::
::: {class='col-md-8'}

![ ](figures/availability.png){ width=70% }
:::
::::

***

## **Total alkalinity vs salinity relationship**
![ ](figures/p_at_s.png)

***

## **Time series of key variables**
![ ](figures/time_series.png)

***

## **Monthly means**
![ ](figures/ts_boxplots.png)

***

## **Consistency**
![ ](figures/pco2_calc_meas.png)

***

## **Water column stratification**
#::: {style="text-align:center" markdown="1"}
  ![ ](figures/deltas.png)
#:::
***

# **Conclusion**

- **First** high-frequency time series of the carbonate system in the Arctic 
- **Positive aspects**: 
  - Excellent support from AWIPEV staff but work overload
  - IPEV support secured for another 4-year period (2019-2022)
- **Logistical difficulties**:
  - Challenging environment; divers needed to sort out issues
  - Difficult and costly to gather a team and organize field trip at short notice
  - Winter sampling essential (once a month this season, more frequently would be more desirable)
  - Demanding for AWIPEV staff
  - IPEV does not cover the cost of analyses required for calibration (ca. 1000 €)